---
title: "Integer Programming"
author: "Namrah"
date: "2023-11-28"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("lpSolveAPI")
# Loading the LP file
data <- read.lp("lin.lp")
print(data)
```

```{r}
# Defining the required workforce for each day
workersNeededPerDay <- matrix(c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday",
                            18,27,22,26,25,21,19), ncol=2, byrow=FALSE)
colnames(workersNeededPerDay) <- c("Day", "RequiredWorkers")
as.table(workersNeededPerDay)

```
*AP's package handling team operates on a five-day work schedule with two consecutive days off. Their base salary stands at $750 per week. Those working on either Saturday or Sunday receive an additional $25 per day. The available shifts and respective compensations for package handlers are:*
```{r}
# Available shifts and corresponding salaries
shiftsAndCompensation <- matrix(c(1,2,3,4,5,6,7,
                             "Sunday and Monday","Monday and Tuesday","Tuesday and Wednesday",
                             "Wednesday and Thursday","Thursday and Friday",
                             "Friday and Saturday","Saturday and Sunday", "$775",
                             "$800","$800","$800","$800","$775","$750"), ncol=3, byrow=FALSE)
colnames(shiftsAndCompensation) <- c("Shift", "DaysOff", "Wage")
as.table(shiftsAndCompensation)

```

```{r}
# Solving the LP model
solve(data)

```
*Valid models yield an outcome of 0.*

```{r}
# Objective Function - Total Cost
totalCost <- get.objective(data)
totalCost
```
*The total cost is $25,675, representing the overall expenses incurred by the company to ensure adequate staffing daily while minimizing labor expenses.*

```{r}
# Workforce availability per day - variable
availableWorkers <- get.variables(data)
availableWorkers
```

```{r}
# Summary of available workers each day
cat("Summary of Available Workers Each Day:\n")
cat("Monday =", availableWorkers[3] + availableWorkers[4] + availableWorkers[5] + availableWorkers[6] + availableWorkers[7], "Workers\n")
cat("Tuesday =", availableWorkers[4] + availableWorkers[5] + availableWorkers[6] + availableWorkers[7] + availableWorkers[1], "Workers\n")
cat("Wednesday =", availableWorkers[5] + availableWorkers[6] + availableWorkers[7] + availableWorkers[1] + availableWorkers[2], "Workers\n")
cat("Thursday =", availableWorkers[6] + availableWorkers[7] + availableWorkers[1] + availableWorkers[2] + availableWorkers[3], "Workers\n")
cat("Friday =", availableWorkers[7] + availableWorkers[1] + availableWorkers[2] + availableWorkers[3] + availableWorkers[4], "Workers\n")
cat("Saturday =", availableWorkers[1] + availableWorkers[2] + availableWorkers[3] + availableWorkers[4] + availableWorkers[5], "Workers\n")
cat("Sunday =", availableWorkers[2] + availableWorkers[3] + availableWorkers[4] + availableWorkers[5] + availableWorkers[6], "Workers\n")

```






